### 2021-09-02 bayes-sampler-prelim.Rmd
added plate diagrams

---
title: "Bayesian sampler working"
output: html_notebook
---

Analyses of the Bayesian sampler data

```{r}
library(tidyverse)
library(purrr)
```

```{r}
## load all data for experiment 1
df_exp1 <- map_dfr(
  paste0("osfstorage-archive/Experiment 1/", list.files("osfstorage-archive/Experiment 1/")), 
  read_csv,
  col_types=cols()
  )

## load all data for experiment 2
df_exp2 <- map_dfr(
  paste0("osfstorage-archive/Experiment 2/", list.files("osfstorage-archive/Experiment 2/")), 
  read_csv,
  col_types=cols()
  )

# sanity check both data
length(unique(df_exp1$ID)) # should be 59 Ps
length(unique(df_exp2$ID)) # should be 84 Ps

```

# Plotting figure 6 (partial)

Quick sanity-check: let's recreate the top half of figure 6. Looks like things are all as expected.

```{r}
df_exp2 %>% 
  filter(block==1) %>% 
  # filter(ID==1, block==2,)
  mutate(
    category = case_when(
      grepl("(windy|cloudy)", querydetail) ~ "{windy, cloudy}",
      grepl("(cold|rainy)", querydetail) ~ "{cold, rainy}",
      TRUE ~ "other"
  ) 
  ) %>% 
    filter(category!="other") %>% 
  mutate(querytype = ordered(querytype,
                             levels = c(
                               "A",
                                "B",
                                "notA",
                                "notB",
                                "AandB",
                                "notAandB",
                                "AandnotB",
                                "notAandnotB",
                                "AorB",
                                "notAorB",
                                "AornotB",
                                "notAornotB",
                                "AgB",
                                "notAgB",
                                "AgnotB",
                                "notAgnotB",
                               "BgA",
                                "notBgA",
                                "BgnotA",
                                "notBgnotA"
                             ))) %>% 
  group_by(querytype, category) %>% 
  mutate(estimate = estimate/100) %>% 
  summarize(
    M = mean(estimate),
    se = sd(estimate)/sqrt(n())
  ) %>% 
  ungroup() %>% 
  ggplot(aes(x=querytype, y=M, ymin=M-se, ymax=M+se)) +
  geom_bar(stat="identity", position="dodge", width = .5, fill="grey50") +
  geom_errorbar(width=.1) +
  facet_wrap(~category, ncol=1) +
  theme_bw(base_size=14) +
  theme(
    panel.grid=element_blank(),
    axis.text.x=element_text(angle=45, hjust=1)
    ) +
  ylim(0,1)
```

bayesian sampler model is written in equation 15 as:

$$E[P_{BS}(A)] = P(A)\frac{N}{N + 2\beta} + \frac{\beta}{N+2\beta}$$

Which we should expect would be Beta distributed, given that it is itself a probability. This has the nice form of a linear regression equation, except that none of the terms in this model are actually observed! The big thing that's missing is $P(A)$. The authors' solution is to treat these as free parameters / unobserved latent variables, but to fix them with a constraint that e.g. in a 4-term conditional probability table the probabilities must sum to one, so there are only 3 degrees of freedom.

I think I can use a simplex variable in Stan to accomplish this.

Meanwhile the PT+N model has a bunch of different equations depending on the type of probability query. For a simple marginal probability it is (per equation 10):

$$E[P_{PT+N}(A)] = (1-2d)P(A)+d$$
Which is likewise a nice linear form but without any variables directly observed. The other equations are given in equations 11-13.

```{r}
# Bayesian sampler model

# for a single person, avg across blocks:
# prior on N
# prior on beta
# prior on N' and beta' (for conjunction and disjunction)
# uninformative prior on unobserved true probabilities (simplex)

# for hierarchical model with many Ps, avg across blocks:
# group-level prior for N, beta, N', beta'
# no hierarchical structure for prior on unobserved true probabilities

# for hierarchical model with many Ps, individual trials:
# share observations across blocks, replace formula

# the actual psychological model is that the sampling can differ each time tho
# so that doesn't seem quite right?
```

## Simulating data

First I'll need to simulate some data. I'll start simulating for one participant, and I'll try to use the format of the final data as much as I can so things translate.

```{r}
# do that

# then fit one real participant

# then expand to hierarchical version

# then fit it all!
```