### 2021-09-02 paper-rmd/ptn_platediagram.tex

\begin{tikzpicture}

     % Define nodes
     \node[obs]                               (y) {$y_{ijk}$};
     \node[det, above=of y] (mu) {$\mu_{ijk}$};
     \node[det, above=of mu, xshift=-1.2cm]  (d) {$d_j$};
     \node[det, above=of mu, xshift=1.2cm]  (d_prime) {$d_j'$};
     \node[latent, above=of d]  (delta) {$\delta_j$};
     \node[latent, above=of d_prime]  (delta_delta) {$\Delta\delta_j$};
     \node[latent, right=2cm of y]  (kappa) {$\kappa$};
     \node[latent, left=1cm of mu]            (p) {$\vec{p_{jk}}$};

     % Connect the nodes
     \edge {mu, kappa} {y} ; %
     \edge {d, d_prime} {mu}
     \edge {delta} {d, d_prime}
     \edge {delta_delta} {d_prime}
     \edge {p} {mu}

     % Plates
     {
       \tikzset{plate caption/.append style={below=15pt of #1.south east}};
       \tikzstyle{plateIJK} = [align=center];
       \plate[inner sep=0.25cm, xshift=-.0625cm] {plateIJK} {(mu)(y)} {$ijk$} ;
    }
     {
       \tikzset{plate caption/.append style={above right=0pt of #1.north east}};
       \plate[inner sep=0.25cm, xshift=-.12cm, yshift=-.12cm] {plateJ} {(mu)(y)(p)(d)(d_prime)(delta)(delta_delta)} {$j$} ;
     }
     {
       \tikzset{plate caption/.append style={above right=0pt and 0pt of #1.north west}};
       \plate[inner sep=0.25cm] {plateIJ} {(p)(mu)(y)} {$jk$};
     }

   %   \plate {} {(mu)(y)(p)(mup.north west)(mup.south west)} {$M$} ;

   \end{tikzpicture}