### 2022-09-16 paper-rmd/bs_platediagram_mlm.tex

\begin{figure}
  \begin{minipage}{.5\textwidth}
    \flushright
    \begin{tikzpicture}
      % Define nodes
      \node[obs]                               (y) {$y_{ijk}$};
      \node[det, above=of y] (mu) {$\mu_{ijk}$};
      \node[det, above=of mu, xshift=-1.2cm]  (d) {$d_j$};
      \node[det, above=of mu, xshift=1.2cm]  (d_prime) {$d_j'$};
      \node[latent, above=of d]  (delta) {$\delta_j$};
      \node[latent, above=of d_prime]  (delta_delta) {$\Delta\delta_j$};
      \node[latent, right=2cm of y]  (kappa) {$\kappa$};
      \node[latent, left=1cm of mu]            (p) {$\vec{p_{jk}}$};
      \node[latent, above=of delta, xshift=-.5cm]  (delta_pop) {$\overline{\delta}$};
      \node[latent, above=of delta_delta, xshift=-.5cm]  (delta_delta_pop) {$\overline{\Delta\delta}$};
      \node[latent, above=of delta, xshift=.5cm]  (sigma_delta) {$\sigma_{\delta}$};
      \node[latent, above=of delta_delta, xshift=.5cm]  (sigma_delta_delta) {$\sigma_{\Delta\delta}$};
      % Connect the nodes
      \edge {mu, kappa} {y} ; %
      \edge {d, d_prime} {mu}
      \edge {delta} {d, d_prime}
      \edge {delta_delta} {d_prime}
      \edge {p} {mu}
      \edge {delta_pop} {delta}
      \edge {delta_delta_pop} {delta_delta}
      \edge {sigma_delta} {delta}
      \edge {sigma_delta_delta} {delta_delta}

      % Plates
      {
        \tikzset{plate caption/.append style={below=15pt of #1.south east}};
        \tikzstyle{plateIJK} = [align=center];
        \plate[inner sep=0.25cm, xshift=-.0625cm] {plateIJK} {(mu)(y)} {$i$} ;
     }
      {
        \tikzset{plate caption/.append style={above right=0pt of #1.north east}};
        \plate[inner sep=0.25cm, xshift=-.12cm, yshift=-.12cm] {plateJ} {(mu)(y)(p)(d)(d_prime)(delta)(delta_delta)} {$j$} ;
      }
      {
        \tikzset{plate caption/.append style={above right=0pt and 0pt of #1.north west}};
        \plate[inner sep=0.25cm] {plateIJ} {(p)(mu)(y)} {$k$};
      }

\end{tikzpicture}
  \end{minipage}%
  \begin{minipage}{.5\textwidth}
    \flushleft
    \small
    \begin{align*}
  y_{ijk} &\sim Beta\big(\mu_{ijk} \kappa, (1-\mu_{ijk}) \kappa \big) \\
  \mu_{ijk} &= f_{BS}(\overrightarrow{p_{jk}}, d_j, d_j')  \\
  d_j &= \frac{1}{3} \ \text{logistic}(\delta_j) \\
  d_j' &= \frac{1}{3} \ \text{logistic}\big(\delta_j + \exp(\Delta\delta_j)\big) \\
  \overrightarrow{p_{jk}} &\sim Dirichlet(\overrightarrow{1}) \\
  \overline{\delta} &\sim Normal(-1, 1) \\
  \overline{\Delta\delta} &\sim Normal(0, .50) \\
  log(\sigma_{\delta}) &\sim Normal(-1, 1) \\
  log(\sigma_{\Delta\delta}) &\sim Normal(-1, 1) \\
  \delta_j &\sim Normal(\overline{\delta}, \sigma_{\delta}) \\
  \Delta\delta_j &\sim Normal(\overline{\Delta\delta}, \sigma_{\Delta\delta}) \\
  \kappa &\sim \mathit{Half \hyph Cauchy(20)} \\
\end{align*}

  \end{minipage}
  \caption{Hierarchical complex Bayesian Sampler model diagram and formula specifications. Circular nodes are parameters, shaded nodes are observations, and squared nodes are deterministic functions of parameters. Plates signify values defined for $i$ trials, $j$ participants, and $k$ conditions.}
\end{figure}