### 2022-09-16 paper-rmd/ptn_platediagram.tex

\begin{figure}
  \begin{minipage}{.5\textwidth}
    \flushright
    \begin{tikzpicture}

     % Define nodes
     \node[obs]                               (y) {$y_{ijk}$};
     \node[det, above=of y] (mu) {$\mu_{ijk}$};
     \node[det, above=of mu, xshift=-1.2cm]  (d) {$d_j$};
     \node[det, above=of mu, xshift=1.2cm]  (d_prime) {$d_j'$};
     \node[obs, right=.75cm of mu]  (x) {$x_{ijk}$};
     \node[latent, above=of d]  (delta) {$\delta_j$};
     \node[latent, above=of d_prime]  (delta_delta) {$\Delta\delta_j$};
     \node[latent, right=2cm of y]  (kappa) {$\kappa$};
     \node[latent, left=.75cm of mu]            (p) {$\vec{\theta_{jk}}$};

     % Connect the nodes
     \edge {mu, kappa} {y} ; %
     \edge {d, d_prime, x, p} {mu}
     \edge {delta} {d, d_prime}
     \edge {delta_delta} {d_prime}

     % Plates
     {
       \tikzset{plate caption/.append style={below=15pt of #1.south east}};
       \tikzstyle{plateIJK} = [align=center];
       \plate[inner sep=0.125cm] {plateIJK} {(mu)(y)(x)} {$i$} ;
    }
     {
       \tikzset{plate caption/.append style={above right=0pt of #1.north east}};
       \plate[inner sep=0.125cm] {plateJ} {(mu)(y)(p)(d)(d_prime)(delta)(delta_delta)(x)} {$j$} ;
     }
     {
       \tikzset{plate caption/.append style={above right=0pt and 0pt of #1.north west}};
       \plate[inner sep=0.125cm] {plateIJ} {(p)(mu)(y)(x)} {$k$};
     }

   %   \plate {} {(mu)(y)(p)(mup.north west)(mup.south west)} {$M$} ;

\end{tikzpicture}
  \end{minipage}%
  \begin{minipage}{.5\textwidth}
    \flushleft
    \small
    \begin{align*}
      y_{ijk} &\sim Beta\big(\mu_{ijk} \kappa, (1-\mu_{ijk}) \kappa \big) \\
      \mu_{ijk} &= f_{PT+N}(\overrightarrow{\theta_{jk}}, x_{ijk}, d_j, d_j)  \\
      d_j &= \frac{1}{2} \ \text{logistic}(\delta_j) \\
      d_j’ &= \frac{1}{2} \ \text{logistic}\big(\delta_j + \exp(\Delta\delta_j)\big) \\
      \overrightarrow{p_{jk}} &\sim Dirichlet(\overrightarrow{1}) \\
      \delta_j &\sim Normal(0,1) \\
      \Delta\delta_j &\sim Normal(0,1) \\
      \kappa &\sim \mathit{Half \hyph Cauchy}(20)
    \end{align*}
  \end{minipage}
  \caption{Complex unpooled PT+N model diagram and formula specifications. Circular nodes are parameters, shaded nodes are observations, and squared nodes are deterministic functions of parameters. Plates signify values defined for $i$ trials, $j$ participants, and $k$ conditions.}
\end{figure}