### 2022-12-14 create-paper-figures.Rmd
updated rmd and avg code

---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
library(patchwork)
library(latex2exp)
library(reticulate)
use_condaenv("bsampler")
```

```{python}
# import numpyro
# numpyro.util.set_host_device_count(4)
# import numpyro.distributions as dist
# from numpyro.infer import MCMC, NUTS, Predictive, DiscreteHMCGibbs

from jax import numpy as jnp
from jax import random

import numpy as np
import pandas as pd
# import seaborn as sns
import arviz as az

# from dfply import *
from siuba import _
import siuba as s

import pickle

from lib.helpers import *
from lib.models import *
from lib.icc import *

```

```{python}
df1 = load_query_avg_data_exp1() # see data_helpers.py
df1 = df1[~df1.querytype.str.contains('or')]
# df1 = df1[df1["condition"]!=2] # filter out "warm/snowy" as per paper
df1.head()
```

```{python}
df2 = load_query_avg_data_exp2() # see data_helpers.py
df2 = df2[df2["condition"]!=2] # filter out "warm/snowy" as per paper
df2.head()
```

```{python}
fits = dict()
fits["exp1"] = dict()
fits["exp2"] = dict()

# fits["exp1"]["Relative Freq."] = pickle.load(open("local/exp1/az_data_rf.p", "rb"))
# fits["exp1"]["Bayesian Sampler simple"] = pickle.load(open("local/exp1/az_data_bs_simple.p", "rb"))
# fits["exp1"]["Bayesian Sampler complex"] = pickle.load(open("local/exp1/az_data_bs.p", "rb"))
# fits["exp1"]["Bayesian Sampler simple hier."] = pickle.load(open("local/exp1/az_data_bs_mlm_simple.p", "rb"))
# fits["exp1"]["Bayesian Sampler complex hier."] = pickle.load(open("local/exp1/az_data_bs_mlm.p", "rb"))
# fits["exp1"]["PT+N complex"] = pickle.load(open("local/exp1/az_data_ptn.p", "rb"))
# fits["exp1"]["PT+N simple"] = pickle.load(open("local/exp1/az_data_ptn_simple.p", "rb"))
# fits["exp1"]["PT+N complex hier."] = pickle.load(open("local/exp1/az_data_ptn_mlm.p", "rb"))
# fits["exp1"]["PT+N simple hier."] = pickle.load(open("local/exp1/az_data_ptn_mlm_simple.p", "rb"))
# fits["exp1"]["Unrestricted BS hier."] = pickle.load(open("local/exp1/az_data_ptn_mlm_simplecond.p", "rb"))
fits["exp1"]["Unrestricted BS hier."] = az.from_netcdf("local/exp1/az_data_ptn_mlm_simplecond.nc")

# fits["exp2"]["Relative Freq."] = pickle.load(open("local/exp2/az_data_rf.p", "rb"))
# fits["exp2"]["Bayesian Sampler simple"] = pickle.load(open("local/exp2/az_data_bs_simple.p", "rb"))
# fits["exp2"]["Bayesian Sampler complex"] = pickle.load(open("local/exp2/az_data_bs.p", "rb"))
# fits["exp2"]["Bayesian Sampler simple hier."] = pickle.load(open("local/exp2/az_data_bs_mlm_simple.p", "rb"))
# fits["exp2"]["Bayesian Sampler complex hier."] = pickle.load(open("local/exp2/az_data_bs_mlm.p", "rb"))
# fits["exp2"]["PT+N complex"] = pickle.load(open("local/exp2/az_data_ptn.p", "rb"))
# fits["exp2"]["PT+N simple"] = pickle.load(open("local/exp2/az_data_ptn_simple.p", "rb"))
# fits["exp2"]["PT+N complex hier."] = pickle.load(open("local/exp2/az_data_ptn_mlm.p", "rb"))
# fits["exp2"]["PT+N simple hier."] = pickle.load(open("local/exp2/az_data_ptn_mlm_simple.p", "rb"))
# fits["exp2"]["Unrestricted BS hier."] = pickle.load(open("local/exp2/az_data_ptn_mlm_simplecond.p", "rb"))
fits["exp2"]["Unrestricted BS hier."] = az.from_netcdf("local/exp2/az_data_ptn_mlm_simplecond.nc")

# loos = dict()
# loos["exp1"] = dict()
# loos["exp2"] = dict()
# 
# loos["exp1"]["Relative Freq."] = pickle.load(open("local/exp1/loo_rf.p", "rb"))
# loos["exp1"]["Bayesian Sampler simple"] = pickle.load(open("local/exp1/loo_bs_simple.p", "rb"))
# loos["exp1"]["Bayesian Sampler complex"] = pickle.load(open("local/exp1/loo_bs.p", "rb"))
# loos["exp1"]["Bayesian Sampler simple hier."] = pickle.load(open("local/exp1/loo_bs_mlm_simple.p", "rb"))
# loos["exp1"]["Bayesian Sampler complex hier."] = pickle.load(open("local/exp1/loo_bs_mlm.p", "rb"))
# loos["exp1"]["PT+N complex"] = pickle.load(open("local/exp1/loo_ptn.p", "rb"))
# loos["exp1"]["PT+N simple"] = pickle.load(open("local/exp1/loo_ptn_simple.p", "rb"))
# loos["exp1"]["PT+N complex hier."] = pickle.load(open("local/exp1/loo_ptn_mlm.p", "rb"))
# loos["exp1"]["PT+N simple hier."] = pickle.load(open("local/exp1/loo_ptn_mlm_simple.p", "rb"))
# loos["exp1"]["Unrestricted BS hier."] = pickle.load(open("local/exp1/loo_ptn_mlm_simplecond.p", "rb"))
# 
# loos["exp2"]["Relative Freq."] = pickle.load(open("local/exp2/loo_rf.p", "rb"))
# loos["exp2"]["Bayesian Sampler simple"] = pickle.load(open("local/exp2/loo_bs_simple.p", "rb"))
# loos["exp2"]["Bayesian Sampler complex"] = pickle.load(open("local/exp2/loo_bs.p", "rb"))
# loos["exp2"]["Bayesian Sampler simple hier."] = pickle.load(open("local/exp2/loo_bs_mlm_simple.p", "rb"))
# loos["exp2"]["Bayesian Sampler complex hier."] = pickle.load(open("local/exp2/loo_bs_mlm.p", "rb"))
# loos["exp2"]["PT+N complex"] = pickle.load(open("local/exp2/loo_ptn.p", "rb"))
# loos["exp2"]["PT+N simple"] = pickle.load(open("local/exp2/loo_ptn_simple.p", "rb"))
# loos["exp2"]["PT+N complex hier."] = pickle.load(open("local/exp2/loo_ptn_mlm.p", "rb"))
# loos["exp2"]["PT+N simple hier."] = pickle.load(open("local/exp2/loo_ptn_mlm_simple.p", "rb"))
# loos["exp2"]["Unrestricted BS hier."] = pickle.load(open("local/exp2/loo_ptn_mlm_simplecond.p", "rb"))
```

```{python}
# exp1_model_comparison = compare(loos["exp1"], method="BB-pseudo-BMA")
# exp2_model_comparison = compare(loos["exp2"], method="BB-pseudo-BMA")

exp1_model_comparison = pd.read_csv("local/exp1/model-comparison.csv")
exp2_model_comparison = pd.read_csv("local/exp2/model-comparison.csv")

exp1_model_comparison = exp1_model_comparison >> s.mutate(Experiment = "Exp. 1", model = exp1_model_comparison["Unnamed: 0"])
exp2_model_comparison = exp2_model_comparison >> s.mutate(Experiment = "Exp. 2", model = exp1_model_comparison["Unnamed: 0"])

model_comparison = exp1_model_comparison.append(exp2_model_comparison)
```

```{r}
py$model_comparison %>% 
  filter(model!="Relative Freq.") %>% 
  mutate(ul=d_loo + dse*2, ll = d_loo - dse*2) %>% 
  ggplot(aes(y=reorder(model,-d_loo), x = d_loo, xmin=ll, xmax=ul)) +
  geom_vline(xintercept=0, linetype="dashed", color="grey") +
  geom_point(size=2) +
  geom_errorbarh(height=.01, size=.5) +
  facet_wrap(~Experiment, scales="free_x") +
  labs(x = TeX("$\\widehat{elpd}_{LOO}$ difference"), y = "Model") +
  theme_bw() +
  theme(panel.grid=element_blank())
```

```{python}
model1 = fits["exp1"]["Unrestricted BS hier."]
model1.posterior["d_prime_pop"] = model1.posterior.d_base_pop + np.exp(model1.posterior.d_delta_pop)
post1 = posterior_to_df_mlm(model1, ["k"], ["d_subj", "d_prime_subj"])

model2 = fits["exp2"]["Unrestricted BS hier."]
model2.posterior["d_prime_pop"] = model2.posterior.d_base_pop + np.exp(model2.posterior.d_delta_pop)
post2 = posterior_to_df_mlm(model2, ["k"], ["d_subj", "d_prime_subj"])

param_posterior1 = model1.posterior.stack(draws=("chain", "draw"))[["d_base_pop", "d_prime_pop"]].to_dataframe()
param_posterior2 = model2.posterior.stack(draws=("chain", "draw"))[["d_base_pop", "d_prime_pop"]].to_dataframe()

param_posterior1 = param_posterior1 >> s.mutate(Experiment = "Exp. 1")
param_posterior2 = param_posterior2 >> s.mutate(Experiment = "Exp. 2")

param_posterior = param_posterior1.append(param_posterior2)
```

```{r}
plt_forest_exp1 <- py$post1 %>% 
  gather(var, val, d_subj, d_prime_subj) %>% 
    group_by(ID, var) %>% 
    summarize(
        M = mean(val),
        ll = quantile(val, .025),
        ul = quantile(val, .975),
    ) %>% 
  mutate(var = if_else(var=="d_subj", "d", "d'")) %>% 
  ggplot(aes(x=reorder(ID,M), y = M, ymin=ll, ymax=ul, color=var)) +
  geom_pointrange(position=position_dodge(width=.5), size=.25) +
  labs(title="Exp. 1", color="Variable", x = "Participant", y = "Estimate") +
  theme_bw() +
  theme(
      panel.grid = element_blank(), 
      legend.position="bottom", 
      axis.text.x = element_text(angle=90, hjust=1, vjust=.5, size=5)
      )

plt_forest_exp2 <- py$post2 %>% 
  gather(var, val, d_subj, d_prime_subj) %>% 
    group_by(ID, var) %>% 
    summarize(
        M = mean(val),
        ll = quantile(val, .025),
        ul = quantile(val, .975),
    ) %>% 
  mutate(var = if_else(var=="d_subj", "d", "d'")) %>% 
  ggplot(aes(x=reorder(ID,M), y = M, ymin=ll, ymax=ul, color=var)) +
  geom_pointrange(position=position_dodge(width=.5),size=.25) +
  labs(title="Exp. 2", color="Variable", x = "Participant", y = "Estimate") +
  theme_bw() +
  theme(
      panel.grid = element_blank(), 
      legend.position="bottom", 
      axis.text.x = element_text(angle=90, hjust=1, vjust=.5, size=5)
      )

(plt_forest_exp1 + coord_fixed(ratio=59/.5)) + 
  (plt_forest_exp2 + coord_fixed(ratio=59/.5)) + 
  plot_layout(guides="collect") & theme(legend.position = 'bottom')
```

```{python}
preds1 = fits["exp1"]["Unrestricted BS hier."].posterior_predictive.mean(dim=['chain', 'draw']).yhat.to_numpy()
preds2 = fits["exp2"]["Unrestricted BS hier."].posterior_predictive.mean(dim=['chain', 'draw']).yhat.to_numpy()
```

```{r}
plt_pp <- py$df1 %>% 
  mutate(pred = py$preds1, Experiment = "Exp. 1") %>% 
  bind_rows(
    py$df2 %>% 
    mutate(pred = py$preds2, Experiment = "Exp. 2")
  ) %>% 
  ggplot(aes(x=pred, y= estimate)) +
  geom_abline(slope=1, intercept=0, linetype="dashed", color="grey") +
  geom_point(alpha=.25, shape=16) +
  facet_wrap(~Experiment) +
  theme_bw() +
  theme(aspect.ratio=1) +
  labs(x="Predicted", y = "Observed") +
  theme(panel.grid=element_blank()) +
  scale_x_continuous(labels=c("0",".25",".50",".75","1")) +
  scale_y_continuous(labels=c("0",".25",".50",".75","1"))

plt_pp
```

```{r}

## calcualte inequalities

plt_ineq_exp2 <- py$df2 %>% 
  mutate(pred = py$preds2) %>% 
  mutate(condition = if_else(condition==0, "{cloudy, windy}", "{cold, rainy}" )) %>% 
  gather(var, val, estimate, pred) %>% 
  select(-querydetail) %>% 
  pivot_wider(names_from = querytype, values_from=val) %>% 
  mutate(
    z1 = A + B - AandB - AorB,
    z2 = A + notAandB - B - AandnotB,
    z3 = A + notAandB - AorB,
    z4 = B + AandnotB - AorB,
    z5 = AandnotB + AandB - A,
    z6 = notAandB + AandB - B,
    z7 = AandnotB + notAandB + AandB - AorB,
    z8 = AandnotB + notAandB + 2*AorB - A -B,
    z9 = AgB*B - BgA*A,
    z10 = AgB*B + AgnotB*notB - A,
    z11 = BgA*A + BgnotA*notA - B,
    z12 = BgA*A + AgnotB*notB - A,
    z13 = AgB*B + BgnotA*notA - B,
    z14 = AgnotB*notB + B - BgnotA*notA - A,
    z15 = AandB - AgB*B,
    z16 = AandB - BgA*A,
    z17 = AandB - A + AgnotB*notB,
    z18 = AandB - B + BgnotA*notA
  ) %>% 
  select(ID, var, condition, starts_with("z")) %>% 
  gather(inequality, val, -ID,-var,-condition) %>% 
  # mutate(inequality = as.numeric(gsub("z","", inequality))) %>% 
  mutate(inequality = ordered(inequality, levels = paste0("z",1:18))) %>% 
  mutate(
    # condition = if_else(condition=="estimate", "Human data", "Model Pred."),
    var = if_else(var=="estimate","Observed", "Predicted")
    ) %>% 
  group_by(var, inequality, condition) %>% 
  summarize(
    M = mean(val),
    SD = sd(val),
    se = sd(val)/sqrt(n())
  ) %>% 
  ggplot(aes(x=inequality, y = M, ymin = M-2*se, ymax=M+2*se, color=var)) +
  geom_pointrange(size=.25) +
  geom_hline(yintercept=0, linetype="dashed") +
  facet_wrap(~condition, ncol=1) +
  theme_bw() +
  theme(panel.grid=element_blank())

```

```{python}
df1_all = load_query_avg_data_exp1()
# model1_all = pickle.load(open("local/exp1/az_data_ptn_mlm_simplecond-all.p", "rb"))
model1_all = az.from_netcdf("local/exp1/az_data_ptn_mlm_simplecond-all.nc")
preds1_all = model1_all.posterior_predictive.mean(dim=['chain', 'draw']).yhat.to_numpy()
```

```{r}
plt_ineq_exp1 <- py$df1_all %>% 
  mutate(pred = py$preds1_all) %>% 
  gather(var, val, estimate, pred) %>% 
  select(-querydetail) %>% 
  pivot_wider(names_from = querytype, values_from=val) %>% 
  mutate(condition = if_else(condition==0, "{icy, frosty}", "{normal, typical}" )) %>% 
  mutate(
    z1 = A + B - AandB - AorB,
    z2 = A + notAandB - B - AandnotB,
    z3 = A + notAandB - AorB,
    z4 = B + AandnotB - AorB,
    z5 = AandnotB + AandB - A,
    z6 = notAandB + AandB - B,
    z7 = AandnotB + notAandB + AandB - AorB,
    z8 = AandnotB + notAandB + 2*AorB - A -B,
    z9 = AgB*B - BgA*A,
    z10 = AgB*B + AgnotB*notB - A,
    z11 = BgA*A + BgnotA*notA - B,
    z12 = BgA*A + AgnotB*notB - A,
    z13 = AgB*B + BgnotA*notA - B,
    z14 = AgnotB*notB + B - BgnotA*notA - A,
    z15 = AandB - AgB*B,
    z16 = AandB - BgA*A,
    z17 = AandB - A + AgnotB*notB,
    z18 = AandB - B + BgnotA*notA
  ) %>% 
  select(ID, var, condition, starts_with("z")) %>% 
  gather(inequality, val, -ID,-var,-condition) %>% 
  # filter(inequality %in% paste0("z",10:13)) %>% 
  # mutate(inequality = as.numeric(gsub("z","", inequality))) %>% 
  mutate(inequality = ordered(inequality, levels = paste0("z",1:18))) %>% 
  mutate(
    # condition = if_else(condition=="estimate", "Human data", "Model Pred."),
    var = if_else(var=="estimate","Observed", "Predicted")
    ) %>% 
  group_by(var, inequality, condition) %>% 
  summarize(
    M = mean(val),
    SD = sd(val),
    se = sd(val)/sqrt(n())
  ) %>% 
  ggplot(aes(x=inequality, y = M, ymin = M-2*se, ymax=M+2*se, color=var)) +
  geom_pointrange(size=.25) +
  geom_hline(yintercept=0, linetype="dashed") +
  facet_wrap(~condition, ncol=1) +
  theme_bw() +
  theme(panel.grid=element_blank())

```

```{r}
# In Experiment 1, like participants' responses,  the model's estimates are very slightly positive for {icy, frosty} and very slightly negative for {normal, typical}. This pattern replicates the qualitative pattern reported by Zhu and colleagues.
plt_ineq_exp1 + labs(title="Exp. 1") + 
  plt_ineq_exp2 + labs(title="Exp. 2") + 
  plot_layout(guides="collect") & 
  theme(legend.position="bottom", axis.text.x = element_text(angle=90, hjust=1, vjust=.5)) & 
  labs(color="", y = "Estimate", x = "Identity")
```

```{python}
# az_ptn_exp1 = pickle.load(open("local/az-ptn_simplecond_mlm_trial_level_disc-exp1svi.p", "rb"))
# az_ptn_exp2 = pickle.load(open("local/az-ptn_simplecond_mlm_trial_level_disc-exp2svi.p", "rb"))

az_ptn_exp1 = az.from_netcdf("local/az-ptn_simplecond_mlm_trial_level_disc-exp1svi.nc") # fix filenames
az_ptn_exp2 = az.from_netcdf("local/az-ptn_simplecond_mlm_trial_level_disc-exp2svi.nc")

posterior_ptn1 = posterior_to_df_mlm(
    az_ptn_exp1,
    ["d_base_pop","d_delta_pop","d_base_sd","d_delta_sd"],
    ["d_subj", "d_prime_subj", "k"]
)

posterior_ptn2 = posterior_to_df_mlm(
    az_ptn_exp2,
    ["d_base_pop","d_delta_pop","d_base_sd","d_delta_sd"],
    ["d_base_r", "d_delta_r", "d_subj", "d_prime_subj", "k"]
)

# posterior_ptn1 = pickle.load(open("local/posterior_ptn1.p", "rb"))
# posterior_ptn2 = pickle.load(open("local/posterior_ptn2.p", "rb"))
```

```{r}
plt_dk_ptn1 <- py$posterior_ptn1 %>% 
    group_by(ID) %>% 
    summarize(
        k_est = mean(k),
        k_ll = quantile(k, .05),
        k_ul = quantile(k, .95),
        d_est = mean(d_subj),
        d_ll = quantile(d_subj, .05),
        d_ul = quantile(d_subj, .95)
    ) %>% 
    filter(k_est < 100) %>%
    # filter(ID!=51, ID!=45, ID!=29, ID!=28) %>%
    ggplot(aes(x=k_est, y=d_est)) +
    geom_point() +
    geom_errorbar(aes(ymin=d_ll, ymax=d_ul), width=0., alpha=.25) +
    geom_errorbarh(aes(xmin=k_ll, xmax=k_ul), height=0, alpha=.25) +
    # geom_smooth(method="lm") +
    labs(y="d", x="N", title="Exp. 1") +
    theme_bw() +
    theme(aspect.ratio=1) +
    theme(panel.grid = element_blank())

plt_dk_ptn2 <- py$posterior_ptn2 %>% 
    group_by(ID) %>% 
    summarize(
        k_est = mean(k),
        k_ll = quantile(k, .05),
        k_ul = quantile(k, .95),
        d_est = mean(d_subj),
        d_ll = quantile(d_subj, .05),
        d_ul = quantile(d_subj, .95)
    ) %>% 
#     s.filter(_.k_est < 50.) >>
    # filter(ID!=51, ID!=45, ID!=29, ID!=28) %>% 
    ggplot(aes(x=k_est, y=d_est)) +
    geom_point() +
    geom_errorbar(aes(ymin=d_ll, ymax=d_ul), width=0., alpha=.25) +
    geom_errorbarh(aes(xmin=k_ll, xmax=k_ul), height=0, alpha=.25) +
    # geom_smooth(method="lm") +
    labs(y="d", x="N", title="Exp. 2") +
    theme_bw() +
    theme(aspect.ratio=1) +
    theme(panel.grid = element_blank())

plt_dprimek_ptn1 <- py$posterior_ptn1 %>% 
    group_by(ID) %>% 
    summarize(
        k_est = mean(k),
        k_ll = quantile(k, .05),
        k_ul = quantile(k, .95),
        d_est = mean(d_prime_subj),
        d_ll = quantile(d_prime_subj, .05),
        d_ul = quantile(d_prime_subj, .95)
    ) %>% 
    filter(k_est < 100) %>%
    # filter(ID!=51, ID!=45, ID!=29, ID!=28) %>%
    ggplot(aes(x=k_est, y=d_est)) +
    geom_point() +
    geom_errorbar(aes(ymin=d_ll, ymax=d_ul), width=0., alpha=.25) +
    geom_errorbarh(aes(xmin=k_ll, xmax=k_ul), height=0, alpha=.25) +
    # geom_smooth(method="lm") +
    labs(y="d'", x="N") +
    theme_bw() +
    theme(aspect.ratio=1) +
    theme(panel.grid = element_blank())

plt_dprimek_ptn2 <- py$posterior_ptn2 %>% 
    group_by(ID) %>% 
    summarize(
        k_est = mean(k),
        k_ll = quantile(k, .05),
        k_ul = quantile(k, .95),
        d_est = mean(d_prime_subj),
        d_ll = quantile(d_prime_subj, .05),
        d_ul = quantile(d_prime_subj, .95)
    ) %>% 
    # s.filter(_.k_est < 50.) %>% 
    # filter(ID!=51, ID!=45, ID!=29, ID!=28) %>% 
    ggplot(aes(x=k_est, y=d_est)) +
    geom_point() +
    geom_errorbar(aes(ymin=d_ll, ymax=d_ul), width=0., alpha=.25) +
    geom_errorbarh(aes(xmin=k_ll, xmax=k_ul), height=0, alpha=.25) +
    # geom_smooth(method="lm") +
    labs(y="d'", x="N") +
    theme_bw() +
    theme(aspect.ratio=1) +
    theme(panel.grid = element_blank())
```

```{r}
plt_dk_ptn1  + plt_dprimek_ptn1 + plt_dk_ptn2 + plt_dprimek_ptn2
```

```{r}
plt_dprimek_ptn1 + geom_smooth(method="lm")
```

```{r}
plt_d_avgmodel <- py$param_posterior %>% 
  mutate(
    d = plogis(d_base_pop)/2,
    d_prime = plogis(d_prime_pop)/2
    ) %>% 
  gather(Variable, val, d, d_prime) %>% 
  mutate(Variable = if_else(Variable=="d", "d", "d'")) %>% 
  ggplot(aes(x=val, fill=Variable, color=Variable)) +
  geom_density(position="identity", alpha=.6) +
  geom_vline(xintercept=.3333, linetype="dashed", color="grey") +
  facet_wrap(~Experiment) +
  xlim(0,.5) +
  labs(title="Participant-level query-average model")

plt_d_trialmodel <- py$posterior_ptn1 %>% 
  mutate(Experiment = "Exp. 1") %>% 
  bind_rows(
    py$posterior_ptn2 %>% 
      mutate(Experiment = "Exp. 2")
  ) %>% 
  filter(ID==0) %>% 
  select(Experiment, draw, d_base_pop, d_base_sd, d_delta_pop, d_delta_sd) %>% 
  mutate(
    d = plogis(d_base_pop)/2,
    d_prime = plogis(d_base_pop + exp(d_delta_pop))/2
    ) %>% 
  gather(Variable, val, d, d_prime) %>% 
  mutate(Variable = if_else(Variable=="d", "d", "d'")) %>% 
  ggplot(aes(x=val, fill=Variable, color=Variable)) +
  geom_density(position="identity", alpha=.6) +
  geom_vline(xintercept=.3333, linetype="dashed", color="grey") +
  facet_wrap(~Experiment) +
  xlim(0,.5) +
  labs(title="Trial-level model")

plt_d_avgmodel + plt_d_trialmodel + plot_layout(ncol=1, guides="collect") & labs(x="Estimate") & theme_bw() & theme(panel.grid=element_blank())
```